From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-19 Pi+/paper/finite.tex
Diagrams

\begin{itemize}
\item swap the $i$-the element with the $(i+1)$-th element, that is, adjacent swaps, or
\item swap the $i$-th element with the $j$-th element, for arbitrary $i$s and $j$s, or
\item swap the $i$-th element with an element at a fixed position, or
\item flip a prefix $\Fin[k]$ of $\Fin[n]$ for $k \leq n$, or
\item cyclically shift any subset of $\Fin[n]$.
\end{itemize}

@@

12. Here, we are choosing a presentation based on the adjacent
transpositions generators - i.e. primitive operations of the form "swap
i-th and i+1 elements". In the language of sorting, this corresponds to
what the insertion sort or bubble sort are doing.
In the explanations below, we will use diagrams to draw the permutations.

12a. There two very obvious laws that such operation has to preserve:
 - Swapping the same two elements two times in a row does not change
anything (cancel)
TODO diagram.
 - When swapping two pairs of elements on a positions far away, it does not
matter in which order one does it (swap). (We can slide the wires).
TODO diagram.
But there is also one more law that these operations have to satisfy, in
the situation when we are swapping adjacent pairs of elements. It is best
seen on a diagram (braid)
TODO diagram.

>>>>

\begin{itemize}
\item swap the $i$-the element with the $(i+1)$-th element, that is, adjacent swaps, or
\item swap the $i$-th element with the $j$-th element, for arbitrary $i$-s and $j$-s, or
\item swap the $i$-th element with an element at a fixed position, or
\item flip a prefix $\Fin[k]$ of $\Fin[n]$ for $k \leq n$, or
\item cyclically shift any subset of $\Fin[n]$.
\end{itemize}

@@

To do the proof, we are using a presentation based on the adjacent
transpositions, i.e. where the primitive operations are adjacent swaps. There
are three laws that these operations have to satisfy - it is easiest to
visualize them on diagrams.
\begin{itemize}
\item Swapping the same two elements two times in a row is the same as doing nothing:
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ s_1, s_1}};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \draw [-stealth](0,0) -- node[anchor=south] {cancel} (2,0);
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={1, 1}};
    \end{tikzpicture}
\item When swapping two pairs of elements on a positions far away, it does not
matter in which order it is done (We can slide the wires).
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ 1, s_1}};
    \end{tikzpicture}
    \hspace{0cm}
    \dots
    \hspace{0cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ s_1, 1 }};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \draw [-stealth](0,0) -- node[anchor=south] {swap} (2,0);
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ s_1, 1}};
    \end{tikzpicture}
    \hspace{0cm}
    \dots
    \hspace{0cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 2,
        thick]
        {braid={ 1, s_1}};
    \end{tikzpicture}

\item seen on a diagram (braid)
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 3,
        thick]
        {braid={ s_2, s_1, s_2}};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \draw [-stealth](0,0) -- node[anchor=south] {braid} (2,0);
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}
        \pic[local bounding box=my braid,braid/.cd, 
        number of strands = 3,
        thick]
        {braid={ s_1, s_2, s_1}};
    \end{tikzpicture}

\end{itemize}

Diagram of long-braid:
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 2,
    thick]
    {braid={1, 1, 1, 1, s_1, 1}};
\end{tikzpicture}
\hspace{0cm}
\dots
\hspace{0cm}
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 5,
    thick]
    {braid={ s_4, s_3, s_2, s_1, 1, s_4}};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
    \draw [-stealth](0,0) -- node[anchor=south] {long-braid} (2,0);
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 2,
    thick]
    {braid={1, 1, 1, 1, 1, s_1}};
\end{tikzpicture}
\hspace{0cm}
\dots
\hspace{0cm}
\begin{tikzpicture}
    \pic[local bounding box=my braid,braid/.cd, 
    number of strands = 5,
    thick]
    {braid={ s_3, s_4, s_3, s_2, s_1, 1}};
\end{tikzpicture}
\hspace{1cm}
